---
title: "ISTA 320 Final Project"
author: "Cole Dryer"
date: "Spring 2022"
output: html_document
---

```{r setup, include=FALSE}
# remember to add any other libraries you need here:
library(knitr)
library(ggplot2)
library(tidyverse)
library(RColorBrewer)
library(ggthemes)
opts_chunk$set(echo = TRUE)
```

This project is about student performance when it comes to high school math. 
(data from www.kaggle.com/datasets/matt40/ordinal-likert-survey-on-hs-math)

# Data Wrangling

```{r}
# ENTER YOUR DATA WRANGLING CODE HERE

df <- read.csv("data/6460_final_data.csv")

df_tidy <- df %>%
  select(`grade_level`:`study_difficulty`,
         `est_current_gpa`)

anxiety_df <- df_tidy %>%
  select(`grade_level`, `exam_anxiety`, `est_current_gpa`)

pivot_df <- df_tidy %>%
  select(`math_conf`:`study_difficulty`) %>%
  pivot_longer(cols = c("math_conf":"study_difficulty"),
               names_to = "factor",
               values_to = "likert rating")

count_df <- pivot_df %>%
  count(`factor`, `likert rating`)

#count_df %>%
  #head()

count_df <- count_df %>%
  mutate(likert_factor = factor(`likert rating`))

count_df <- count_df %>%
  mutate(n = ifelse(`likert rating` < 3,
                    -n, n))
```

# Data Visualization

```{r}

#PLOT 1

anxiety_df %>%
  ggplot(aes(x=est_current_gpa, y=exam_anxiety, fill=grade_level)) +
  geom_col() +
  labs(y="Anxiety Level", x="Estimated GPA", fill="Grade",
       title="Student Anxiety to GPA Bar Plot",
       caption = "data from www.kaggle.com/datasets/matt40/ordinal-likert-survey-on-hs-math")

```

This plot shows which students are showing the most anxiety. The colors represent the different high school grade levels and each bar represents a GPA score. You can see that the higher the GPA a student has actually somewhat leads to more anxiety, according to this plot. 

```{r}

#PLOT 2

count_df %>%
  ggplot(aes(x = n,
             y = reorder(`factor`, n),
             fill = `likert_factor`)) +
  geom_col(data = count_df %>% filter(n > 0),
           position = position_stack(reverse = TRUE)) +
  geom_col(data = count_df %>% filter(n < 1),
           position = position_stack(reverse = FALSE)) +
scale_fill_manual(values = brewer.pal(6,"BrBG")) +
  theme_linedraw() +
labs(y = "",
       fill = "Rating",
       title = "Student Performance Likert Survey: Visual Factor Bar Plot",
       subtitle = "Students’ satisfaction ratings (1-5) for different 
factors of school access, performance, and emotional ratings",
       caption = "data from www.kaggle.com/datasets/matt40/ordinal-likert-survey-on-hs-math")
```

In this plot above, I created a visual representation of different likert ratings from students and how they scored different aspects of the high school, when it comes to math. Using this likert scale and a color theme for the bars, it is easy to see which factors/aspects need improvement and how students are feeling in general when it comes to high school math.

```{r}

#PLOT 3

mean_df <- pivot_df %>%
  group_by(`factor`) %>%
  summarize(avg_rating = mean(`likert rating`, na.rm = TRUE)) %>%
  ungroup()

#mean_df%>%
  #head()

mean_df %>%
  ggplot(aes(x = avg_rating,
             y = `factor`,
             group = 1)) +
  geom_point() +
  geom_line() +
  labs(y = "",
       x = "Average Likert Rating",
       title = "Student Performance Likert Survey: Mean Ratings Line Plot",
       subtitle = "Average likert ratings from students",
       caption = "data from www.kaggle.com/datasets/matt40/ordinal-likert-survey-on-hs-math")
```

In this plot above, I created a line plot showing the average student likert ratings for each service provided by the high school. The line helps show the difference in each average as well as which factors/aspects need the most improvement.




